---
title: "MILES"
author: "Franziska Lang"
date: "`r Sys.Date()`"
output: 
  html_document:
    code_folding: hide
    toc: true
    toc_depth: 3    
    toc_float: true
    highlight: pygments
editor_options: 
  chunk_output_type: console
knit: (
  function(inputFile, encoding) { 
    rmarkdown::render(
      inputFile, 
      encoding = encoding, 
      output_file = file.path(dirname(inputFile), "overview_miles.html")) }
      )
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE, echo = F)

date_today <- gsub("-", "", Sys.Date())
library(tidyverse)
library(ggpubr)
library(gridExtra)
library(gtools)

#ADJUST THIS PATH!
setwd("./miles_publication/")


source("data_analysis/functions_evaluation_models.R")
theme_set(theme_bw())


col_entity <- c("#E7CDC2", "#80A0C7", "#394165")

path_plots <- "plots/"

export_plots <- F

export_table <- F
```


## AIM

Summary of MILES performance in context of the mutation type and mutation entity FOR THE ANALYSIS WITHOUT PATIENTS WITH STABLE DISEASE. 
Import Tuning results and calculate median ROC-AUC over 10x10 CV to determine the best hyper-parameter set that will represent performance of MILES on the respective data subset. 

## 1 Import results from hyperparameter tuning in nested CV

### 1.1 MEL + RCC

#### 1.1.1 SNVs

```{r , echo=FALSE, message=FALSE, warning = FALSE, results=FALSE}


path_to_files_snvs <- "data_for_publication/MILES_results/SNV/miles_all_wo_SD/"


dat_snv_import <- import_cv(path_to_files_snvs)
dat_snv <- transform_performance(dat_snv_import)
best_model_snv_rcc_mel <- get_best_model(dat_snv)


```

#### 1.1.2 fusion genes


```{r , echo=FALSE, message=FALSE, warning = FALSE, results=FALSE}


path_to_files_fusions <- "data_for_publication/MILES_results/fusion_gene//miles_all_wo_SD/"


dat_fusion_import <- import_cv(path_to_files_fusions)
dat_fusion <- transform_performance(dat_fusion_import)
best_model_fusion_rcc_mel <- get_best_model(dat_fusion)


```


#### 1.1.3 combined


```{r , echo=FALSE, message=FALSE, warning = FALSE, results=FALSE}


path_to_files_combined <- "data_for_publication/MILES_results/combined///miles_all_wo_SD/"


dat_combined_import <- import_cv(path_to_files_combined)
dat_combined <- transform_performance(dat_combined_import)
best_model_combined_rcc_mel <- get_best_model(dat_combined)

```



### 1.1.4 INDEL

```{r , echo=FALSE, message=FALSE, warning = FALSE, results=FALSE}


path_to_files_indel <-  "data_for_publication/MILES_results/INDEL/miles_all_wo_SD/"


dat_indel_import <- import_cv(path_to_files_indel)
dat_indel <- transform_performance(dat_indel_import)
best_model_indel_rcc_mel <- get_best_model(dat_indel)


```


### 1.2 MEL

#### 1.2.1 SNVs

```{r , echo=FALSE, message=FALSE, warning = FALSE, results=FALSE}


path_to_files_snvs <- "data_for_publication/MILES_results/SNV/miles_mel_wo_SD/"


dat_snv_import <- import_cv(path_to_files_snvs)
dat_snv <- transform_performance(dat_snv_import)
best_model_snv_mel <- get_best_model(dat_snv)


```

#### 1.2.2 combined


```{r , echo=FALSE, message=FALSE, warning = FALSE, results=FALSE}


path_to_files_combined <- "data_for_publication/MILES_results/combined//miles_mel_wo_SD/"


dat_combined_import <- import_cv(path_to_files_combined)
dat_combined <- transform_performance(dat_combined_import)
best_model_combined_mel <- get_best_model(dat_combined)


```


#### 1.2.3 Fusion genes

```{r , echo=FALSE, message=FALSE, warning = FALSE, results=FALSE}


path_to_files_fusions <- "data_for_publication/MILES_results/fusion_gene/miles_mel_wo_SD/"


dat_fusion_import <- import_cv(path_to_files_fusions)
dat_fusion <- transform_performance(dat_fusion_import)
best_model_fusion_mel <- get_best_model(dat_fusion)

```



### 1.2.4 INDEL

```{r , echo=FALSE, message=FALSE, warning = FALSE, results=FALSE}


path_to_files_indel <- "data_for_publication/MILES_results/INDEL/miles_mel_wo_SD/"


dat_indel_import <- import_cv(path_to_files_indel)
dat_indel <- transform_performance(dat_indel_import)
best_model_indel_mel <- get_best_model(dat_indel)


```

### 1.3 RCC

#### 1.3.1 SNVs

```{r , echo=FALSE, message=FALSE, warning = FALSE, results=FALSE}


path_to_files_snvs <- "data_for_publication/MILES_results/SNV/miles_rcc_wo_SD/"


dat_snv_import <- import_cv(path_to_files_snvs)
dat_snv <- transform_performance(dat_snv_import)
best_model_snv_rcc <- get_best_model(dat_snv)



```

#### 1.3.2 combined


```{r , echo=FALSE, message=FALSE, warning = FALSE, results=FALSE}


path_to_files_combined <- "data_for_publication/MILES_results/combined//miles_rcc_wo_SD/"


dat_combined_import <- import_cv(path_to_files_combined)
dat_combined <- transform_performance(dat_combined_import)
best_model_combined_rcc <- get_best_model(dat_combined)

```


#### 1.3.3 Fusion genes

```{r , echo=FALSE, message=FALSE, warning = FALSE, results=FALSE}


path_to_files_fusions <- "data_for_publication/MILES_results/fusion_gene///miles_rcc_wo_SD/"


dat_fusion_import <- import_cv(path_to_files_fusions)
dat_fusion <- transform_performance(dat_fusion_import)
best_model_fusion_rcc <- get_best_model(dat_fusion)


```



### 1.3.4 INDEL

```{r , echo=FALSE, message=FALSE, warning = FALSE, results=FALSE}


path_to_files_indel <- "data_for_publication/MILES_results/INDEL/miles_rcc_wo_SD/"


dat_indel_import <- import_cv(path_to_files_indel)
dat_indel <- transform_performance(dat_indel_import)
best_model_indel_rcc <- get_best_model(dat_indel)

```


## 2 Identification of best models


```{r , echo=FALSE, message=FALSE, warning = FALSE, results=FALSE}

# summary of qualitative model 
auc_rcc_mel <- list(SNV = best_model_snv_rcc_mel, "Fusion gene" = best_model_fusion_rcc_mel, "INDEL" = best_model_indel_rcc_mel, "combined" = best_model_combined_rcc_mel)
auc_rcc_mel <- lapply(auc_rcc_mel, function(x) x%>% mutate(sigma2 = as.numeric(sigma2), c = as.numeric(c)))
auc_rcc_mel <- bind_rows(auc_rcc_mel, .id = "class")


auc_mel <- list(SNV = best_model_snv_mel, "Fusion gene" = best_model_fusion_mel, "INDEL" = best_model_indel_mel, "combined" = best_model_combined_mel)
auc_mel <- lapply(auc_mel, function(x) x%>% mutate(sigma2 = as.numeric(sigma2), c = as.numeric(c)))
auc_mel <- bind_rows(auc_mel, .id = "class")

auc_rcc <- list(SNV = best_model_snv_rcc, "Fusion gene" = best_model_fusion_rcc, "INDEL" = best_model_indel_rcc, "combined" = best_model_combined_rcc)
auc_rcc <- lapply(auc_rcc, function(x) x%>% mutate(sigma2 = as.numeric(sigma2), c = as.numeric(c)))
auc_rcc <- bind_rows(auc_rcc, .id = "class")

# identify best models per group
auc_overall <- bind_rows(list("MEL+RCC" = auc_rcc_mel, "MEL" = auc_mel, "RCC" = auc_rcc), .id = "entity")

paht_entity = c("MEL+RCC" = "miles_all_wo_SD", "MEL" = "miles_mel_wo_SD", "RCC" = "miles_rcc_wo_SD" )
auc_overall_path <- auc_overall %>%
  mutate(mut_path = ifelse(class == "Fusion gene", "fusion_gene", class)) %>%
  mutate(entity_path = paht_entity[entity]) %>%
  mutate(path_file = paste0("data_for_publication/MILES_results/",mut_path, "/", entity_path, "/miles_", gsub(" ", "", format(sigma2, scientific = FALSE)), "_", c, ".tsv"))
  


if(export_table){
  nam <- "data_for_publication/MILES_results/MILES_best_models_without_SD.tsv"
  write_delim(auc_overall_path, nam, delim = "\t")
}




```


