---
title: "Comparison of quantiative and qualitative analysis"
author: "Franziska Lang"
date: "`r Sys.Date()`"
output: 
  html_document:
    code_folding: hide
    toc: true
    toc_depth: 3    
    toc_float: true
    highlight: pygments
editor_options: 
  chunk_output_type: console
knit: (
  function(inputFile, encoding) { 
    rmarkdown::render(
      inputFile, 
      encoding = encoding, 
      output_file = file.path(dirname(inputFile), "icb_quant_qual_data.html")) }
      )
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE, echo = F)


#ADJUST THIS PATH!
setwd("./miles_publication/")


date_today <- gsub("-", "", Sys.Date())
library(tidyverse)
library(ggpubr)
library(gridExtra)
library("wesanderson")
source("data_analysis/functions_icb2neo.R")
theme_set(theme_bw())


col_mutation <- wes_palette("FantasticFox1",3)
col_expressed <- wes_palette("Chevalier1",2)
col_entity <- c("#E7CDC2", "#80A0C7", "#394165")

path_to_plots <- "plots/"
export_plots <- FALSE

export_table <- FALSE

```

```{setup, include=FALSE}


# technical issues with iCaM --> SNVs + INDELs will be counted as NA 
# Hugo Pt 8 
# McDermott Pt164370145747331
# 


# no SNVs/INDELS were detected for the following patients why here we can count 0
# Riaz pre Pt24
# McDermott Pt204886393976620

```



## AIM

The aim is to compare the quantitative and qualitative analysis 

## 1 With patients with stable disease



```{r , echo=FALSE, message=FALSE, warning = FALSE, results=FALSE}

file_quant <- "data_for_publication/MILES_results/20220308_quantitative_model_mean_auc_nested_cv.txt"
file_qual <- "data_for_publication/MILES_results/20220308_performance_models_with_SD.tsv"

dat_quant <- read_delim(file_quant)
dat_qual <- read_delim(file_qual)

dat_qual <- dat_qual %>%
  select(entity, class, auc, auc_iqr)

# these cols do not contain mean and sd but median +/- iqr
dat_quant <- dat_quant %>%
  dplyr::rename(auc = mean_auc) %>%
  dplyr::rename(auc_iqr = sd_auc) %>%
  select(-num)


dat <- bind_rows(list(qualitative = dat_qual, quantitative = dat_quant), .id = "model")

dat_mel_combined <- dat %>%
  filter(entity == "MEL" & class == "combined")

dat_rcc_fus <- dat %>%
  filter(entity == "MEL+RCC" & class == "Fusion gene")
```

```{r , echo=FALSE, message=FALSE, warning = FALSE, results=FALSE}

p1 <- dat_mel_combined %>%
  ggplot(aes(model, auc ))+
  geom_bar(stat = "identity", position = "dodge")+
  geom_hline(yintercept = 0.5) +
  geom_errorbar(aes(ymin=auc-auc_iqr , ymax=auc+auc_iqr ), width=.2,
                 position=position_dodge(.9), color = "darkgrey") +
  scale_fill_manual(values = rev(col_entity))+
  ggtitle("MEL : combined")
 

p2 <- dat_rcc_fus %>%
  ggplot(aes(model, auc ))+
  geom_bar(stat = "identity", position = "dodge")+
  geom_hline(yintercept = 0.5) +
  geom_errorbar(aes(ymin=auc-auc_iqr , ymax=auc+auc_iqr ), width=.2,
                 position=position_dodge(.9), color = "darkgrey") +
  scale_fill_manual(values = rev(col_entity)) + 
  ggtitle("MEL+RCC : fusion genes")

if(export_plots){
  nam <- paste0(path_to_plots, "/", date_today, "_comparison_quant_qual_with_SD.pdf")
  pdf(nam)
}

ggarrange(p1, p2,
          labels = c("C", "D"),
          ncol = 2, nrow = 2)  

if(export_plots){
  dev.off()
}

```


The boxplot shows for each cohort the predicted candidate count per patient in respect to the mutation class (expressed and non-expressed are considered).  

```{r , echo=FALSE, message=FALSE, warning = FALSE, results=FALSE}

# tall-skinny data frame of all cohorts
# NOTE: NA is counted as 0 --> while this holds true for most of the patients-mut class combination there were a very few patients for which NA should derive from a MyMut error that could not be solved
DF2 <- transfrom_count2tall(numbers_all)
DF2_plot <- DF2 %>%
  filter(!class == "all")


## FIGURE 2A
if(export_plots){
  nam <- paste0(path_to_plots, "/", date_today, "_boxplot_number_all_candidates_class.pdf")
  pdf(nam, width = 10, height = 6)
}

xlabs <- paste(levels(DF2_plot$cohort),"\n(n=",table(DF2_plot$cohort[which(DF2_plot$class == "SNV")]),")",sep="")

DF2_plot %>%
  ggplot(aes(x = cohort, y = count + 1, fill = class)) +
  geom_boxplot(outlier.fill = NULL, outlier.shape = 21) +
  scale_y_continuous(trans ='log10') +
  scale_fill_manual(values = col_mutation) + 
  scale_color_manual(values =col_mutation) +
  theme(panel.grid.major = element_blank())+
  scale_x_discrete(labels=xlabs)


if(export_plots){
  dev.off()
  #ggsave(nam, p,  width=6, height=10)
}  


```

Proportions of mutation source for neoantigen candidates per patient:  

```{r , echo=FALSE, message=FALSE, warning = FALSE, results=FALSE}
 
# substitute NAs with 0 in the main data frame and calculate percentage 
number_summary <- lapply(names(numbers_all), function(x){
  technical_problems <- c("hugo" = "Pt8", "miao" = "PtRCC_106","mcdermott_Atezo" =  "Pt164370145747331", "riaz_pre" = "no", "vanallen" = "no")
  numbers_all[[x]] %>%
    mutate(snv = ifelse(is.na(snv) & patientIdentifier != technical_problems[x], 0, snv )) %>%
    mutate(INDEL = ifelse(is.na(INDEL) & patientIdentifier != technical_problems[x], 0, INDEL )) %>%
    mutate(all = rowSums(.[2:4], na.rm = FALSE))%>%
    mutate(percentage_snv = snv/all, 
           percentage_indel = INDEL/all,
           percentage_fusion = `fusion gene`/all) %>%
    filter(!is.na(percentage_snv))
  
})
names(number_summary) <- names(number_snvs)

## FIGURE 2B
if(export_plots){
  nam <- paste0(path_to_plots, "/", date_today, "_proportion_neoantigen_class.pdf")
  pdf(nam)
}
par(mar=c(10,4,4,4))
lapply(names(number_summary), function(x){

  make_proportion_barplot(number_summary[[x]], main = x)
})
if(export_plots){
  dev.off()
}    
```


The following boxplot show the number all predicted neoantigen candidates per patient and the number of expressed neoantigen candidates for SNVs, INDELs and fusion genes.  


```{r , echo=FALSE, message=FALSE, warning = FALSE, results=FALSE}

# USE THIS AS SUPPLEMENTAL FIGURE
DF2_expressed <- transfrom_count2tall(numbers_expressed)
DF2_expressed_plot <- DF2_expressed %>%
  filter(class != "all")
DF_all_expressed <- bind_rows(list(all = DF2_plot, expressed = DF2_expressed_plot), .id = "category")

# compare numbers of all predicted and expressed neoantigen candidates
if(export_plots){
  nam <- paste0(path_to_plots, "/", date_today, "_comparison_all_expressed.pdf")
  pdf(nam)
}

boxplot_all_selected(DF_all_expressed, main = "all")
boxplot_all_selected_facet(DF_all_expressed)

if(export_plots){
  dev.off()
}  

```


The following boxplot show the number all predicted neoantigen candidates per patient and the neoantigen candidates that generate at least one MHC binding epitope for SNVs, INDELs and fusion genes.  

```{r , echo=FALSE, message=FALSE, warning = FALSE, results=FALSE}

# USE THIS AS SUPPLEMENTAL FIGURE
DF2_binding <- transfrom_count2tall(numbers_binding)
DF2_binding_plot <- DF2_binding %>%
  filter(class != "all")
DF_all_binding <- bind_rows(list(all = DF2_plot, expressed = DF2_binding_plot), .id = "category")

# compare numbers of all predicted and binding neoantigen candidates

if(export_plots){
  nam <- paste0(path_to_plots, "/", date_today, "_comparison_all_binding.pdf")
  pdf(nam)
}

boxplot_all_selected(DF_all_binding, main = "all")
boxplot_all_selected_facet(DF_all_binding)
if(export_plots){
  dev.off()
}  

```


### 1.2 Neoantigen candidate quantity and response to ICB

The following boxplots compare the neoantigen candidate load (without any filter) between responding and non-responding patients. 

```{r , echo=FALSE, message=FALSE, warning = FALSE, results=FALSE}

#cohorts_with_rna <- c("hugo", "vanallen", "riaz_pre", "riaz_on", "mcdermott_Atezo", "mcdermott_Atezo+Bev")
cohorts_with_rna <- c("hugo", "vanallen", "riaz_pre",  "mcdermott_Atezo")

# prepare a response data-frame
response_snv <- lapply(dat_transformed,function(x){
  x%>%
    select(patientIdentifier, binary_response) %>%
    distinct
})
response_fusions <- lapply(dat_fusions_transformed,function(x){
  x%>%
    select(patientIdentifier, binary_response) %>%
    distinct
})

response_df <- response_snv
df_resp <- lapply(names(response_snv), function(cohort){
  if(cohort %in% cohorts_with_rna){
    res <- response_snv[[cohort]] %>%
    full_join(response_fusions[[cohort]], by = "patientIdentifier")%>%
    mutate(binary_response = ifelse(is.na(binary_response.x), binary_response.y , ifelse(is.na(binary_response.y), binary_response.x, binary_response.x) )) %>%
      select(patientIdentifier, binary_response)
  }else{
    response_snv[[cohort]]%>%
      select(patientIdentifier, binary_response)
  }
})

names(df_resp) <- names(response_snv)

df_resp_df <- bind_rows(df_resp, .id = "cohort")

nice_names <- c("Van Allen","Hugo", "Riaz", "Miao", "McDermott")
names(nice_names) <- c( "vanallen" ,"hugo"  ,"riaz_pre" ,"miao" ,"mcdermott_Atezo")
df_resp_df <- df_resp_df %>%
   mutate(cohort = nice_names[cohort]) %>%
  mutate(cohort_pat = paste0(cohort,"_" ,patientIdentifier)) 


```

```{r , echo=FALSE, message=FALSE, warning = FALSE, results=FALSE}

DF2_resp <- DF2 %>%
  mutate(cohort_pat = paste0(cohort,"_" ,patientIdentifier)) %>%
  left_join(df_resp_df, by = "cohort_pat")

```

```{r , echo=FALSE, message=FALSE, warning = FALSE, results=FALSE}

# boxplot showing number of neoantigen candidates load between responder and non-responder
col_response <- c(no=rgb(12,44,132, 255, maxColorValue = 255),
                  yes = rgb(178,34,34, 250 , maxColorValue = 255))

if(export_plots){
  nam <- paste0(path_to_plots, "/", date_today, "_boxplot_response_number_all_candidates.pdf")
  pdf(nam)
}    

DF2_resp %>%
  make_boxplot_response_facet()

if(export_plots){
  dev.off()
}    

pval_all <- compare_means(count ~ binary_response, data = DF2_resp, group.by = c("class", "cohort.x"))
pval_all_all_cohorts <- compare_means(count ~ binary_response, data = DF2_resp, group.by = c("class"))
pval_all_all_cohorts$cohort.x <- "all"

pval_all <- bind_rows(pval_all, pval_all_all_cohorts)

```



### 1.3 Expressed neoantigen candidate quantity and response to ICB

The following boxplots compare the load with expressed number of neoantigen candidates between responding and non-responding patients. 

```{r , echo=FALSE, message=FALSE, warning = FALSE, results=FALSE}

# merge response and number of neoantigen candidates
DF2_expresed_resp <- DF2_expressed %>%
  mutate(cohort_pat = paste0(cohort,"_" ,patientIdentifier)) %>%
  left_join(df_resp_df, by = "cohort_pat")

pval_expressed <- compare_means(count ~ binary_response, data = DF2_expresed_resp, group.by = c("class", "cohort.x"))
pval_expressed_all_cohorts <- compare_means(count ~ binary_response, data = DF2_expresed_resp, group.by = c("class"))
pval_expressed_all_cohorts$cohort.x <- "all"

pval_expressed <- bind_rows(pval_expressed, pval_expressed_all_cohorts)


```

```{r , echo=FALSE, message=FALSE, warning = FALSE, results=FALSE}


if(export_plots){
  nam <- paste0(path_to_plots, "/", date_today, "_boxplot_response_number_expressed_candidates.pdf")
  pdf(nam)
}

DF2_expresed_resp %>%
  make_boxplot_response_facet()

if(export_plots){
  dev.off()
}    


```


### 1.3 Number of neoantigen candidatese meeting defined thresholds and the response to ICB

Here, the neoantigen candidate load is estimated by considering only candidates that harbour MHC I or MHC II binding capabilities meeting a given threshold. 

```{r , echo=FALSE, message=FALSE, warning = FALSE, results=FALSE}

# neoantigen candidate load at different MHC biniding thresholds 

feature1 = "Best_affinity_MHCI_score"
feature2 = "Best_affinity_MHCII_score"

count1000 <- neoag_qual_class_count(dat_transformed, dat_indel_transformed, 
                            dat_fusions_transformed, feature1 = feature1, feature2 = feature2, thres = 1000)
pval1000 <- neoag_qual_class_pval(count1000)
count500 <- neoag_qual_class_count(dat_transformed, dat_indel_transformed, 
                            dat_fusions_transformed, feature1 = feature1, feature2 = feature2, thres = 500)
pval500 <- neoag_qual_class_pval(count500)
count50 <- neoag_qual_class_count(dat_transformed, dat_indel_transformed, 
                            dat_fusions_transformed, feature1 = feature1, feature2 = feature2, thres = 50)
pval50 <- neoag_qual_class_pval(count50)

```




```{r , echo=FALSE, message=FALSE, warning = FALSE, results=FALSE}

# count expressed neoantigen candidates with best MHC I/II affinity < 50, 500, 1000 nM

dat_transformed_expressed <- lapply(dat_pat_expressed, bind_rows, .id = "cohort")
dat_indel_transformed_expressed <- lapply(dat_indel_pat_expressed, bind_rows, .id = "cohort")
dat_fusions_transformed_expressed <- lapply(dat_fusions_pat_expressed, bind_rows, .id = "cohort")

# MHC < 1000
count1000_expressed <- neoag_qual_class_count(dat_transformed_expressed, dat_indel_transformed_expressed, 
                            dat_fusions_transformed_expressed, feature1 = feature1, feature2 = feature2, thres = 1000)
pval1000_expressed <- neoag_qual_class_pval(count1000_expressed)
# MHC < 500
count500_expressed <- neoag_qual_class_count(dat_transformed_expressed, dat_indel_transformed_expressed, 
                            dat_fusions_transformed_expressed, feature1 = feature1, feature2 = feature2, thres = 500)
pval500_expressed <- neoag_qual_class_pval(count500_expressed)
# MHC < 50
count50_expressed <- neoag_qual_class_count(dat_transformed_expressed, dat_indel_transformed_expressed, 
                            dat_fusions_transformed_expressed, feature1 = feature1, feature2 = feature2, thres = 50)
pval50_expressed <- neoag_qual_class_pval(count50_expressed)



```

```{r , echo=FALSE, message=FALSE, warning = FALSE, results=FALSE}
pval_all <- pval_all %>%
  dplyr::rename(cohort = cohort.x)
pval_expressed <- pval_expressed %>%
  dplyr::rename(cohort = cohort.x)


pvals_list <- bind_rows(list(all = pval_all, all_expressed = pval_expressed, 
                             MHC1000 = pval1000, MHC500 = pval500, MHC50 =pval50,
                             MHC1000_expressed = pval1000_expressed, MHC500_expressed = pval500_expressed, MHC50_expressed =pval50_expressed ), .id = "method")

col_pal <- c( RColorBrewer::brewer.pal(5, "GnBu")[3:5], "grey90") 

  

## Supplement Figure 1


S1<- pvals_list %>%
  mutate(p.signif = factor(p.signif)) %>%
  mutate(method = factor(method, levels = rev(c("all", "all_expressed", "MHC1000",  "MHC1000_expressed", "MHC500","MHC500_expressed", "MHC50", "MHC50_expressed")))) %>%
  ggplot(aes(cohort, method, color = p.signif)) +
  geom_point(alpha = 1 , size = 8) + 
  facet_wrap(vars(class)) + 
  scale_color_manual(values =col_pal)+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) 

if(export_plots){
  nam <- paste0(path_to_plots, "/", date_today, "_ICB_dot_plot_pvals_count_methods.pdf")
  pdf(nam, height = 6, width = 10)
}
S1 
if(export_plots){
  dev.off()
}


```


```{r , echo=FALSE, message=FALSE, warning = FALSE, results=FALSE}

# some boxplot to validate finding from dot plot 
if(export_plots){
  nam <- paste0(path_to_plots, "/", date_today, "_ICB_selected_count_response_correl.pdf")
  pdf(nam)
}

# ALL --> SNV 
p1 <- DF2_resp %>% 
  filter(class == "SNV") %>%
  make_boxplot_response(main = "all cohorts : SNV")

p2 <- count50 %>%
  filter(class == "SNV") %>%
  make_boxplot_response(main = "all cohorts : SNV : MHC < 50 ")
  
p3<- count50_expressed %>%
  filter(class == "SNV") %>%
  make_boxplot_response(main = "all cohorts : SNV : expressed + MHC < 50")



# ALL --> INDEL 
p4 <- count50 %>%
  filter(class == "INDEL") %>%
  make_boxplot_response(main = "all : INDEL : MHC < 50 ")
  
p5 <- count50_expressed %>%
  filter(class == "INDEL") %>%
  make_boxplot_response(main = "all : INDEL :  expressed + MHC < 50")

# RIAZ INDEL 
p6 <- count50 %>%
  filter(cohort == "riaz_pre" & class == "INDEL") %>%
  mutate(binary_response = factor(binary_response)) %>%
  make_boxplot_response(main = "Riaz : INDEL : MHC < 50")

p7 <- count50_expressed   %>%
  filter(cohort == "riaz_pre" & class == "INDEL") %>%
  mutate(binary_response = factor(binary_response)) %>%
  make_boxplot_response(main = "Riaz : INDEL : expressed + MHC < 50 ")

ggarrange(p1, p2, p3, p4, p5, p6, p7 , common.legend = TRUE,
          labels = c("C", "D", "E", "F", "G", "H", "I"),
          ncol = 3, nrow = 3)  

if(export_plots){
  dev.off()
}
```


## 2 Density Plots of neoantigen features 

```{r , echo=FALSE, message=FALSE, warning = FALSE, results=FALSE}

features.nummeric <- features_full

plot_density_mutation <- function(feature, snv, indel, fusion) {
  
  class_rename <- c("snv" = "SNV", "indel"= "INDEL", "fusion"= "Fusion gene")
  
  tsnv <-
    tibble(val = snv[[feature]],
           class = "snv")
  tindel <-
    tibble(val = indel[[feature]],
           class = "indel")
  tfusion <-
    tibble(val = fusion[[feature]],
           class = "fusion")
  
  t <- bind_rows(tsnv, tindel, tfusion)
  t <- t %>%
    mutate(class = class_rename[class]) %>%
    mutate(class = factor(class, levels = c("SNV", "INDEL", "Fusion gene")))
  
  ggplot(t, aes(val, fill = class)) + geom_density(alpha = 0.9) + scale_x_log10() +  xlab(feature) + scale_fill_manual(values=col_mutation)+
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())
  
}

pls <- lapply(features.nummeric, function(x){
  plot_density_mutation(feature = x, snv = dat_transformed_comb$all_cohorts, indel = dat_indel_transformed_comb$all_cohorts,
                      fusion = dat_fusions_transformed_comb$all_cohorts)
})

if(export_plots){
  nam <- paste0(path_to_plots, "/", date_today, "_density_features.pdf")
  #pdf(nam)
  ggsave(nam, marrangeGrob(grobs = pls, nrow=2, ncol=2))
  #dev.off()
}else{
  pls
}


```


## 2 ROC-AUC for quantity

To compare performance between qualitative and quantitative approach, 10x10 cross validation is performed for the neoantigen candidate load as a predictor of the response to ICB 

### 2.1 RCC + MEL

#### 2.1.1 SNV

```{r , echo=FALSE, message=FALSE, warning = FALSE, results=FALSE}

# SNVS
DF2_resp_snv <- DF2_resp %>%
  filter(class == "SNV") %>%
  filter(!is.na(binary_response))%>%
  filter(!is.na(count))


DF2_resp_snv_folds <- nested_CV(DF2_resp_snv)
predictions_full <- lapply(DF2_resp_snv_folds, function(y) y$count)
binary_response <- lapply(DF2_resp_snv_folds, function(y) y$binary_response)
auc_snvs <- roc_with_cv(predictions_full = predictions_full, labels = binary_response)


```


#### 2.1.2 Combined

```{r , echo=FALSE, message=FALSE, warning = FALSE, results=FALSE}

DF2_resp_combined <- DF2_resp %>%
  filter(!is.na(binary_response))%>%
  filter(!is.na(count))

DF2_resp_combined_folds <- nested_CV(DF2_resp_combined)
predictions_full <- lapply(DF2_resp_combined_folds, function(y) y$count)
binary_response <- lapply(DF2_resp_combined_folds, function(y) y$binary_response)
auc_combined <- roc_with_cv(predictions_full = predictions_full, labels = binary_response)


```


#### 2.1.3 Fusions

```{r , echo=FALSE, message=FALSE, warning = FALSE, results=FALSE}

DF2_resp_fusions <- DF2_resp %>%
  filter(class == "Fusion gene") %>%
  filter(!is.na(binary_response))%>%
  filter(!is.na(count))

DF2_resp_fusions_folds <- nested_CV(DF2_resp_fusions)
predictions_full <- lapply(DF2_resp_fusions_folds, function(y) y$count)
binary_response <- lapply(DF2_resp_fusions_folds, function(y) y$binary_response)
auc_fusions <- roc_with_cv(predictions_full = predictions_full, labels = binary_response)


```


#### 2.1.4 INDELS

```{r , echo=FALSE, message=FALSE, warning = FALSE, results=FALSE}

DF2_resp_indel <- DF2_resp %>%
  filter(class == "INDEL") %>%
  filter(!is.na(binary_response))%>%
  filter(!is.na(count))

DF2_resp_indel_folds <- nested_CV(DF2_resp_indel)
predictions_full <- lapply(DF2_resp_indel_folds, function(y) y$count)
binary_response <- lapply(DF2_resp_indel_folds, function(y) y$binary_response)
auc_indel <- roc_with_cv(predictions_full = predictions_full, labels = binary_response)


```


### 2.2 MEL

#### 2.2.1 SNV

```{r , echo=FALSE, message=FALSE, warning = FALSE, results=FALSE}

# SNVS
DF2_resp_snv <- DF2_resp %>%
  filter(cohort.x %in% c("Van Allen","Hugo", "Riaz" ))%>%
  filter(class == "SNV") %>%
  filter(!is.na(binary_response))%>%
  filter(!is.na(count))


DF2_resp_snv_folds <- nested_CV(DF2_resp_snv)
predictions_full <- lapply(DF2_resp_snv_folds, function(y) y$count)
binary_response <- lapply(DF2_resp_snv_folds, function(y) y$binary_response)
auc_snvs_mel <- roc_with_cv(predictions_full = predictions_full, labels = binary_response)


```

#### 2.2.2 Combined

```{r , echo=FALSE, message=FALSE, warning = FALSE, results=FALSE}

DF2_resp_combined <- DF2_resp %>%
  filter(cohort.x %in% c("Van Allen","Hugo", "Riaz" ))%>%
  filter(!is.na(binary_response))%>%
  filter(!is.na(count))

DF2_resp_combined_folds <- nested_CV(DF2_resp_combined)
predictions_full <- lapply(DF2_resp_combined_folds, function(y) y$count)
binary_response <- lapply(DF2_resp_combined_folds, function(y) y$binary_response)
auc_combined_mel <- roc_with_cv(predictions_full = predictions_full, labels = binary_response)

```


#### 2.2.3 Fusions

```{r , echo=FALSE, message=FALSE, warning = FALSE, results=FALSE}

DF2_resp_fusions <- DF2_resp %>%
  filter(cohort.x %in% c("Van Allen","Hugo", "Riaz" ))%>%
  filter(class == "Fusion gene") %>%
  filter(!is.na(binary_response))%>%
  filter(!is.na(count))

DF2_resp_fusions_folds <- nested_CV(DF2_resp_fusions)
predictions_full <- lapply(DF2_resp_fusions_folds, function(y) y$count)
binary_response <- lapply(DF2_resp_fusions_folds, function(y) y$binary_response)
auc_fusions_mel <- roc_with_cv(predictions_full = predictions_full, labels = binary_response)


```


#### 2.2.4 INDELS

```{r , echo=FALSE, message=FALSE, warning = FALSE, results=FALSE}

DF2_resp_indel <- DF2_resp %>%
  filter(cohort.x %in% c("Van Allen","Hugo", "Riaz" ))%>%
  filter(class == "INDEL") %>%
  filter(!is.na(binary_response))%>%
  filter(!is.na(count))

DF2_resp_indel_folds <- nested_CV(DF2_resp_indel)
predictions_full <- lapply(DF2_resp_indel_folds, function(y) y$count)
binary_response <- lapply(DF2_resp_indel_folds, function(y) y$binary_response)
auc_indel_mel <- roc_with_cv(predictions_full = predictions_full, labels = binary_response)


```


### 2.3 RCC
#### 2.3.1 SNV

```{r , echo=FALSE, message=FALSE, warning = FALSE, results=FALSE}

# SNVS
DF2_resp_snv <- DF2_resp %>%
  filter(cohort.x %in% c("Miao","McDermott" ))%>%
  filter(class == "SNV") %>%
  filter(!is.na(binary_response))%>%
  filter(!is.na(count))


DF2_resp_snv_folds <- nested_CV(DF2_resp_snv)
predictions_full <- lapply(DF2_resp_snv_folds, function(y) y$count)
binary_response <- lapply(DF2_resp_snv_folds, function(y) y$binary_response)
auc_snvs_rcc <- roc_with_cv(predictions_full = predictions_full, labels = binary_response)


```

#### 2.3.2 Combined

```{r , echo=FALSE, message=FALSE, warning = FALSE, results=FALSE}

# SNVS
DF2_resp_combined <- DF2_resp %>%
  filter(cohort.x %in% c("Miao","McDermott" ))%>%
  filter(!is.na(binary_response))%>%
  filter(!is.na(count))

DF2_resp_combined_folds <- nested_CV(DF2_resp_combined)
predictions_full <- lapply(DF2_resp_combined_folds, function(y) y$count)
binary_response <- lapply(DF2_resp_combined_folds, function(y) y$binary_response)
auc_combined_rcc <- roc_with_cv(predictions_full = predictions_full, labels = binary_response)


```


#### 2.3.3 Fusions

```{r , echo=FALSE, message=FALSE, warning = FALSE, results=FALSE}

DF2_resp_fusions <- DF2_resp %>%
  filter(cohort.x %in% c("Miao","McDermott" ))%>%
  filter(class == "Fusion gene") %>%
  filter(!is.na(binary_response))%>%
  filter(!is.na(count))

DF2_resp_fusions_folds <- nested_CV(DF2_resp_fusions)
predictions_full <- lapply(DF2_resp_fusions_folds, function(y) y$count)
binary_response <- lapply(DF2_resp_fusions_folds, function(y) y$binary_response)
auc_fusions_rcc <- roc_with_cv(predictions_full = predictions_full, labels = binary_response)


```



#### 2.3.4 INDELS

```{r , echo=FALSE, message=FALSE, warning = FALSE, results=FALSE}

DF2_resp_indel <- DF2_resp %>%
  filter(cohort.x %in% c("Miao","McDermott" ))%>%
  filter(class == "INDEL") %>%
  filter(!is.na(binary_response))%>%
  filter(!is.na(count))

DF2_resp_indel_folds <- nested_CV(DF2_resp_indel)
predictions_full <- lapply(DF2_resp_indel_folds, function(y) y$count)
binary_response <- lapply(DF2_resp_indel_folds, function(y) y$binary_response)
auc_indel_rcc <- roc_with_cv(predictions_full = predictions_full, labels = binary_response)


```


#### 2.4 SUMMARY

```{r , echo=FALSE, message=FALSE, warning = FALSE, results=FALSE}

# summary of quantitative model 
auc_rcc_mel <- bind_rows(list(SNV = auc_snvs, "Fusion gene" = auc_fusions, "INDEL" = auc_indel, "combined" = auc_combined), .id = "class")
auc_mel <- bind_rows(list(SNV = auc_snvs_mel, "Fusion gene" = auc_fusions_mel, "INDEL" = auc_indel_mel, "combined" = auc_combined_mel), .id = "class")
auc_rcc <- bind_rows(list(SNV = auc_snvs_rcc, "Fusion gene" = auc_fusions_rcc, "INDEL" = auc_indel_rcc, "combined" = auc_combined_rcc), .id = "class")
auc_overall <- bind_rows(list("MEL+RCC" = auc_rcc_mel, "MEL" = auc_mel, "RCC" = auc_rcc), .id = "entity")



if(export_plots){
  nam <- paste0(path_to_plots, "/", date_today, "_ICB_quantitative_model_bar_auc_nested_cv_with_SD.pdf")
  pdf(nam, height = 6, width = 10)
}
auc_overall %>%
  mutate(entity = factor(entity, levels = c("MEL+RCC", "MEL", "RCC"))) %>%
  ggplot(aes(class, mean_auc, fill = entity))+
  geom_bar(stat = "identity", position = "dodge")+
  geom_hline(yintercept = 0.5) +
  geom_errorbar(aes(ymin=mean_auc-sd_auc , ymax=mean_auc+sd_auc ), width=.2,
                 position=position_dodge(.9), color = "darkgrey") +
  scale_fill_manual(values = rev(col_entity))

if(export_plots){
  dev.off()
}


if(export_table){
  nam <- paste0("data_for_publication/",date_today, "_quantitative_model_mean_auc_nested_cv.txt")
  write_delim(auc_overall, nam)
}

```

#### 3 Overview response distribution 

```{r , echo=FALSE, message=FALSE, warning = FALSE, results=FALSE}

dat_indel_transformed$miao$response <- dat_indel_transformed$miao$best_RECIST

dat_indel_transformed2 <- lapply(dat_indel_transformed , function(x){
  x %>% distinct(patientIdentifier, .keep_all = TRUE) %>%
    select(patientIdentifier, response)
})

df_response <- bind_rows(dat_indel_transformed2, .id = "cohort")


if(export_plots){
  nam <- paste0(path_to_plots, "/", date_today, "_distribution_responses_cohorts.pdf")
  pdf(nam, height = 6, width = 10)
}
df_response %>% count(cohort, response) %>%
  mutate(response = factor(response, levels = rev(c("CR", "PR", "SD", "PD")))) %>%
  mutate(cohort = factor(cohort, levels = c("hugo", "riaz_pre", "vanallen", "mcdermott_Atezo", "miao"))) %>%
  ggplot(aes(cohort,n,  fill = response))+
  scale_fill_manual(values = wes_palette("Zissou1", 4)) +
  geom_bar(stat = "identity")
if(export_plots){
  dev.off()
}
```
